Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
Fit in 8 pages

@@

\section{Introduction}
%Similarities between RL and SL
On a very general level, artificial intelligence addresses the problem of an agent that must select the right actions to solve a task. The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal. This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels). However, the two approaches share many challenges and tools. The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or directly a policy defining the agent's behaviour. Furthermore, when the tasks are characterized by large or continuous state-action spaces, RL needs the powerful function approximators (\eg neural networks) that are the main subject of study of SL.
% SL needs SVRG so RL
In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$. The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums that approximate the expected performance over an unknown data distribution. When generalization to the complete dataset is not taken into consideration, we talk about Empirical Risk Minimization (ERM). Even in this case, stochastic optimization is often used for reasons of efficiency. The idea of stochastic gradient (SG) ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction. At the level of the single iteration, this can be much less expensive than taking into account all the data. However, the sub-sampling of data is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of solving this trade-off, with significant results both in theory and practice. Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems.

@@

\section{Preliminaries}\label{sec:pre}
In this section, we provide the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

\subsection{Policy Gradient}\label{subsec:PolicyGradient}
A Reinforcement Learning task~\citep{sutton1998reinforcement} can be modelled with a discrete-time continuous Markov Decision Process (MDP) $M = \{\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
% and $R$ is the maximum absolute-value reward;
$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
The agent's behaviour is modelled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
% We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
% In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable, but the optimal policy can reach the target state (\ie absorbing state) in at most $H$ steps. This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$. A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
We denote with $p(\tau|\pi)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories (see Appendix~\ref{A:gradient_estimators} for the definition), and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
%
$\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).$
%
Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim p(\cdot|\pi)]{\Reward(\tau)|M}$.
Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.

@@

\subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}z_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $z_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $z_i(\vtheta) = z(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
%% \end{align*}
A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}. 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{See Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

@@

Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
= \Var\left[\gradApp{\vtheta^*}{N}\right].
Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.

@@

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
We mainly leverage on the recent analysis of non-concave SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In Appendix~\ref{app:gauss} we show how Gaussian policies satisfy these assumptions.

@@

\section{Remarks on SVRPG}
The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
Additionally, we provide a variant of SVRPG exploiting a variance-reduction technique for importance weights.
Despite lacking theoretical guarantees, we will show in Section~\ref{sec:exp} that this method can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).

\subsection{Full Gradient Update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\wt{\vtheta}^s}$ feels like a waste of data (the term $\sum_i g(\tau_i) - \omega(\tau_i) g(\tau_i) =0$ since $\vtheta_0 = \wt{\vtheta}$).
In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\wt{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \wt{\vtheta}^s + \alpha\gradApp{\wt{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha 
%        v_t^{s+1}
        \blacktriangledown J(\vtheta^{s+1}_t)
        \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following, we will always assume this practical variant is employed.

\subsection{Meta-Parameter Selection}\label{sec:stopping}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency, while
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
the epoch length $m$ influences the variance introduced by the importance weights. Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations). High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights. We propose to adapt the two meta-parameters jointly.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

@@

\subsection{Normalized Importance Sampling}\label{sec:prac}
As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization~\citep[\eg][]{precup2000eligibility,owenmcbook}.
This technique can reduce the variance of the importance weights at the cost of introducing some bias~\citep[][Chapter 9]{owenmcbook}.
Whether the trade-off is advantageous depends on the specific task.  
Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + \frac{1}{B} \sum_{i=0}^{B-1} \left[g(\tau_i|\vtheta_t)\right]\\ 
                                          &\qquad{} - \frac{1}{\Omega} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right].
\end{align*}
where $\Omega = \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})$.
In Section \ref{sec:exp} we show that self-normalization can provide a performance improvement.

\section{Related Work}
Despite the considerable interest received in SL, variance-reduced gradient approaches have not attracted the RL community.
As far as we know, there are just two applications of SVRG in RL.
The first approach~\citep{du2017svrgpe} aims to exploit SVRG for policy evaluation.
The policy evaluation problem is more straightforward than the one faced in this paper (control problem).
In particular, since the goal is to evaluate just the performance of a predefined policy, the optimization problem is stationary.
The setting considered in the paper is the one of policy evaluation by minimizing the empirical mean squared projected Bellman error (MSPBE) with a linear approximation of the value function. \citet{du2017svrgpe} shown that this problem can be equivalently reformulated as a convex-concave saddle-point problem that is characterized by a finite-sum structure.
This problem can be solved using a variant of SVRG~\citep{Palaniappan2016svrgsaddle} for which convergence guarantees have been provided.
The second approach~\citep{xu2017svrgtrpo} uses SVRG as a practical method to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration. This is just a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
It is worth to mention that, for practical reasons, the authors proposed to use a Newton conjugate gradient method with SVRG.

>>>>

\makeatletter
\g@addto@macro\normalsize{%
 \setlength\abovedisplayskip{6pt}
 \setlength\belowdisplayskip{6pt}
}
\makeatother

@@

\vspace{-0.05in}
\section{Introduction}
\vspace{-0.05in}
%Similarities between RL and SL
On a very general level, artificial intelligence addresses the problem of an agent that must select the right actions to solve a task. The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal. This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels). However, the two approaches share many challenges and tools. The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or directly a policy defining the agent's behaviour. Furthermore, when the tasks are characterized by large or continuous state-action spaces, RL needs the powerful function approximators (\eg neural networks) that are the main subject of study of SL.
% SL needs SVRG so RL
In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$. The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums that approximate the expected performance over an unknown data distribution. When generalization to the complete dataset is not taken into consideration, we talk about Empirical Risk Minimization (ERM). Even in this case, stochastic optimization is often used for reasons of efficiency. The idea of stochastic gradient (SG) ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction. At the level of the single iteration, this can be much less expensive than taking into account all the data. However, the sub-sampling of data is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of solving this trade-off, with significant results both in theory and practice. Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems.

@@

\vspace{-0.05in}
\section{Preliminaries}\label{sec:pre}
\vspace{-0.05in}
In this section, we provide the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

\vspace{-0.05in}
\subsection{Policy Gradient}\label{subsec:PolicyGradient}
\vspace{-0.05in}
A Reinforcement Learning task~\citep{sutton1998reinforcement} can be modelled with a discrete-time continuous Markov Decision Process (MDP) $M = \{\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\}$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$;
% and $R$ is the maximum absolute-value reward;
$\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
The agent's behaviour is modelled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
% We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
% In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
We consider episodic MDPs with effective horizon $H$.\footnote{The episode duration is a random variable, but the optimal policy can reach the target state (\ie absorbing state) in at most $H$ steps. This has not to be confused with a finite horizon problem where the optimal policy is non-stationary.} In this setting, we can limit our attention to trajectories of length $H$. A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
We denote with $p(\tau|\pi)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories (see Appendix~\ref{A:gradient_estimators} for the definition), and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
%
$\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).$
%
Policies can be ranked based on their expected total reward: $J(\pi) = \EVV[\tau \sim p(\cdot|\pi)]{\Reward(\tau)|M}$.
Solving an MDP $M$ means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.

@@

\vspace{-0.05in}
\subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
\vspace{-0.05in}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}z_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $z_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $z_i(\vtheta) = z(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
%% \end{align*}
A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
Under this hypothesis, full gradient (FG) ascent ~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate \cite{nesterov2013introductory}.
However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}. 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.

@@

\vspace{-0.05in}
\section{SVRG in Reinforcement Learning}\label{sec:alg}
\vspace{-0.05in}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{See Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

@@

%\begin{proof}
%Claim (\ref{eq:unbiased}) is from the unbiasedness of the estimators:
%\EVV[]{\gradBlack{\vtheta}} &= \EVV[]{\gradApp{\wt{\vtheta}}{N}}  + \EVV[]{\gradApp{\vtheta}{B}} \\
%&\qquad- \EVV[]{\frac{1}{B}\sum_{i=0}^{B-1}\omega(\tau_i|\vtheta, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} \\
%&= \gradJ{\wt{\vtheta}} + \gradJ{\vtheta} - \gradJ{\wt{\vtheta}} = \gradJ{\vtheta}.
%As for claim (\ref{eq:zerovar}), note that as $\vtheta\to\vtheta^*$, also $\wt{\vtheta}\to\vtheta^*$. Hence, by continuity of $J(\vtheta)$:
%\Var\left[\gradBlack{\vtheta}\right] &\to \Var\left[\gradApp{\vtheta^*}{N}\right] + \frac{1}{B}\Var\left[g(\tau|\vtheta^*)\right. \\
%&\left.- \cancel{\omega(\tau|\vtheta^*,\vtheta^*)}g(\tau|\vtheta^*)\right]
%= \Var\left[\gradApp{\vtheta^*}{N}\right].
%Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.
%\end{proof}

@@

\vspace{-0.05in}
\section{Convergence Guarantees of SVRPG}\label{sec:conv}
\vspace{-0.05in}
In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
We mainly leverage on the recent analysis of non-concave SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In Appendix~\ref{app:gauss} we show how Gaussian policies satisfy these assumptions.

@@

\vspace{-0.05in}
\section{Remarks on SVRPG}
\vspace{-0.05in}
The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
Additionally, we provide a variant of SVRPG exploiting a variance-reduction technique for importance weights.
Despite lacking theoretical guarantees, we will show in Section~\ref{sec:exp} that this method can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).

\vspace{-0.05in}
\subsection{Full Gradient Update}
\vspace{-0.05in}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\wt{\vtheta}^s}$ feels like a waste of data (the term $\sum_i g(\tau_i) - \omega(\tau_i) g(\tau_i) =0$ since $\vtheta_0 = \wt{\vtheta}$).
In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\wt{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \wt{\vtheta}^s + \alpha\gradApp{\wt{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha 
%        v_t^{s+1}
        \blacktriangledown J(\vtheta^{s+1}_t)
        \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following, we will always assume this practical variant is employed.

\vspace{-0.05in}
\subsection{Meta-Parameter Selection}\label{sec:stopping}
\vspace{-0.05in}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency, while
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
the epoch length $m$ influences the variance introduced by the importance weights. Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations). High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights. We propose to adapt the two meta-parameters jointly.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

@@

\vspace{-0.05in}
\subsection{Normalized Importance Sampling}\label{sec:prac}
\vspace{-0.05in}
As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization~\citep[\eg][]{precup2000eligibility,owenmcbook}.
This technique can reduce the variance of the importance weights at the cost of introducing some bias~\citep[][Chapter 9]{owenmcbook}.
Whether the trade-off is advantageous depends on the specific task.  
Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + \frac{1}{B} \sum_{i=0}^{B-1} \left[g(\tau_i|\vtheta_t)\right]\\ 
                                          &\qquad{} - \frac{1}{\Omega} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right].
\end{align*}
where $\Omega = \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})$.
In Section \ref{sec:exp} we show that self-normalization can provide a performance improvement.

\vspace{-0.05in}
\section{Related Work}
\vspace{-0.05in}
Despite the considerable interest received in SL, variance-reduced gradient approaches have not attracted the RL community.
As far as we know, there are just two applications of SVRG in RL.
The first approach~\citep{du2017svrgpe} aims to exploit SVRG for policy evaluation.
The policy evaluation problem is more straightforward than the one faced in this paper (control problem).
In particular, since the goal is to evaluate just the performance of a predefined policy, the optimization problem is stationary.
The setting considered in the paper is the one of policy evaluation by minimizing the empirical mean squared projected Bellman error (MSPBE) with a linear approximation of the value function. \citet{du2017svrgpe} shown that this problem can be equivalently reformulated as a convex-concave saddle-point problem that is characterized by a finite-sum structure.
This problem can be solved using a variant of SVRG~\citep{Palaniappan2016svrgsaddle} for which convergence guarantees have been provided.
The second approach~\citep{xu2017svrgtrpo} uses SVRG as a practical method to solve the optimization problem faced by Trust Region Policy Optimization (TRPO) at each iteration. This is just a direct application of SVRG to a problem having finite-sum structure since no specific structure of the RL problem is exploited.
It is worth to mention that, for practical reasons, the authors proposed to use a Newton conjugate gradient method with SVRG.
